\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2014-01-27 exps.tex
Updated the paper

We have carried out a number of experiments on a set of applications
with real-time constraints. The characteristics of the benchmark set
including required real-time delays and program sizes are shown in
Table~\ref{fig:benchlist}.
{\renewcommand{\arraystretch}{0.9}
%\renewcommand{\texttt}[1] {\begin{lstlisting}[language=Java,morekeywords={delay}] #1 \end{lstlisting}}
\begin{table}[h!]
	\small
	\centering
	\caption{An overview of the benchmark programs}
	\label{fig:benchlist}
	\scalebox{0.9}{
		\begin{tabular}{llp{0.5cm}p{1cm}}
\toprule
\emph{Benchmark}       & \emph{Delay constructs used in each CD} & \emph{\# of CDs} & 
\emph{Lines of code}        
\\ \midrule
HRTCS                  & \texttt{delay(50.3..200.3 ms)} ($CD1$) & 2                  & 56
\\ \midrule
\multirow{5}{*}{Motor} & \texttt{delay(2.4..2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
											 & \texttt{delay(1.667..1.667 ms)}      &                    &                      \\ 
											 & \texttt{delay(0.05..0.05 ms)}       &                    &                      \\ 
											 & \texttt{delay(0.3..0.3 ms)}        &                    &                      \\ 
											 & \texttt{delay(0.733..0.733 ms)}      &                    &
\\ \midrule
\multirow{3}{*}{Printer}
& \texttt{delay(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
& \texttt{delay(1.2..30.3 ms)} ($CD2$)		  &														&											\\
& \texttt{delay(5.6..100.2 ms)} ($CD2$)     &														&											
\\ \midrule
\multirow{2}{*}{AECS} & \texttt{delay(10000 ms)} ($CD1$) 
& \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
& \texttt{delay(10000..10000 ms)} ($CD2$) & &
\\ \bottomrule
\end{tabular}
}
\end{table}
}
{\renewcommand{\arraystretch}{0.6}
\begin{table*}[t!]
\centering
	\caption{Actual delays obtained for \texttt{delay} constructs in the
benchmark programs based on their BCRT and WCRT}
\begin{tabular}{ c c l l l l l l }
	\toprule
	\multicolumn{2}{c}{} & \emph{HRTCS} & \emph{Motor} & \emph{Printer/CD1} &
	\emph{Printer/CD2} & \emph{AECS/CD1} &	\emph{AECS/CD2} \\ 
	\midrule
	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0.0396 ms	& 0.0425 ms & 0.0190 ms &
	0.0263 ms & 0.1784 ms & 0.1284 ms\\ 
	\cmidrule(r){1-8}
	\multicolumn{2}{c}{\emph{WCRT}} &0.1152 ms & 0.1096 ms & 0.0647 ms
	& 0.0491 ms & 0.3476 ms & 0.2991 ms\\ 
	\cmidrule(r){1-8}

	\multirow{2}{*}{Delay 1} & M..N & 50.3-200.3 ms	& 2.4-6.2493 ms	  &
	2.7-10.6 ms & 1.2-30.3 ms & 10000-19484.3481 ms& 10000-23278.2752 ms \\
	\cmidrule(r){2-8}
	& d &1272-1739 & 57 & 143-163 & 46-617 & 56062 &77844\\ 
	\cmidrule(r){1-8}
	\multirow{2}{*}{Delay 2} & M..N & N/A & 1.668-4.3855 ms & N/A &
	5.6-100.2 ms & N/A &10000-23278.2752 ms\\ 
	\cmidrule(r){2-8}
	& d & N/A & 40 & N/A & 213-2041 & N/A &53160\\ 
	\cmidrule(r){1-8}
	\multirow{2}{*}{Delay 3} & M..N & N/A & 0.05-0.2193 ms & N/A & N/A & N/A
	&N/A\\ 
	\cmidrule(r){2-8}
	& d & N/A & 2 & N/A & N/A & N/A &N/A\\ 
	\cmidrule(r){1-8}
	\multirow{2}{*}{Delay 4} & M..N & N/A & 0.3-0.8771 ms & N/A & N/A & N/A
	&N/A\\ 
	\cmidrule(r){2-8}
	& d & N/A & 8 & N/A & N/A & N/A &N/A\\ 
	\cmidrule(r){1-8}
	\multirow{2}{*}{Delay 5} & M..N & N/A & 0.733-1.9735 ms & N/A & N/A & N/A
	&N/A\\ 
	\cmidrule(r){2-8}
	& d & N/A & 18 & N/A
	& N/A & N/A &N/A\\ \bottomrule
\end{tabular}
\label{fig:comparison}
\end{table*}
}

HRTCS is the human response system described in
Section~\ref{sec:motivating-example}. AECS (Access and Environmental
Control System) is an application that controls an intelligent room
environment described in~\cite{aecs_ispa}. The system interacts with two
external timers through input/output signals in order to measure how
long the entrance door to the room is being opened to detect intruders.
The program has been modified such that the delay constructs are used
instead of external timers as shown in Figure~\ref{delaycode}. In the
original AECS, the system needed to trigger external timer by emitting
the output signal in order to wait for a specific amount of time (e.g.
\texttt{TimerTriggered} in Figure~\ref{delaycode:a}). When the timer
expires, it notifies the system through the input signal
\texttt{TimeOut}, which is polled using the \texttt{await}
statement. With the new delay constructs, these timers are not necessary
and delays can be expressed directly in the code as shown in
Figure~\ref{delaycode:b}.

\begin{figure}[h!]
	\vspace{-10pt}
	\centering
  \begin{SubFloat}{\label{delaycode:a}Original program}
      \scriptsize
		\begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,delay,suspend}]
emit TimerTriggered;
trap(T){
  abort(DoorOpened){
    await(TimeOut);
    exit(T); 
  }
}
\end{lstlisting}
  \end{SubFloat}
  \hspace{1cm}%
  \begin{SubFloat}{\label{delaycode:b}Program with delay constructs}
		\begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,delay,suspend}]
// No emit statement
trap(T){
  abort(DoorOpened){
    delay(10000 ms);
    exit(T);  
  }
}
\end{lstlisting}
	\end{SubFloat}
\caption{Replacing external timer with delay construct}
\label{delaycode}
\end{figure}

Motor is a stepper motor controller from~\cite{Bourke2009a}, used for
producing monochrome images on paper by correct application of current
to a print head (a row of resistors). In this example, duration of the
current applied is controlled by the feedback logic using real-time
delays, which are expressed using reactive constructs. Printer is an
example from~\cite{Schneider:1999:CRT:555233}, which models a printer
and a spooler in timed CSP. Its operation can be described as follows:
as soon as the spooler receives a print job it passes it to the printer
through a channel where actual printing process take place. This example
along with HRTCS demonstrates the use of non-deterministic delays.

@@

HRTCS consists of 2 CDs, but only one with real-time delays. AECS also
has 2 CDs; CD1 consist of a single delay statement, while CD2 has 2
delay statements. In this case, as explained previously, delay
statements replace the external timers and corresponding interface (I/O)
signals. The system triggers an alarm 10 seconds after detection of an
intruder or generates a beep sound when an entrance door is being opened
for the same amount of time. In the Printer example, spooler and printer
processes are mapped to separate CDs. Spooler CD sends a job to the
Printer CD between 2.7 to 10.6 ms after receiving it from the
environment (e.g. a user). On the other hand, Printer CD notifies the
environment that it starts printing between 1.2 to 30.3 ms once it
receives the job from the Spooler CD. Actual print job is started
between 5.6 to 100.2 ms after the notification. In this example, bounded
delay ensures that both; channel communication and print job should
occur within the specified amount of time. Lastly, Motor is a
synchronous program with 5 delay statements that control the amount of
current applied to the stepper motor coils.

@@

Table~\ref{fig:comparison} shows the result of our algorithm applied to
each benchmark program. Note that we allowed the compiler to relax the
upper real-time bounds even for bounded (\texttt{M..N}) or exact
(\texttt{M..M}) delays when it failed to find a feasible solution.
However, a user can prohibit such behavior using compiler switch as
described in Section~\ref{sec:tool-chain-flow}.

As a first step, BCRT and WCRT were calculated for all examples. Given
the real-time \mbox{\texttt{delay (M..N)}} provided by programmers,
required logical delay \textit{d} was determined by applying
Algorithm~\ref{alg:1}. In HRTCS, real-time delay of the range
50.3..200.3 (ms) is only satisfied when \emph{d} is between 1272 and
1739 (as shown previously in Section~\ref{sec:find-logic-delay}). Delay
2.7..10.6 (ms) is satisfied in Printer/CD1 with \emph{d} of 143..163.
Similarly, Printer/CD2 satisfies real-time ranges of 1.2..30.3 (ms) and
5.6..100.2 (ms) when \emph{d} is 46..617 and 213..2041, respectively. It
is then the compiler's job to statically choose one of the values in
this range.

@@

The previous results show that for HRTCS and Printer that the compiler
was able to find \textit{d} without relaxing upper bound of
\texttt{delay}. Figure~\ref{fig:graph1} shows the accuracy of our
results by comparing differences between programmer specified upper
bounds (Table~\ref{fig:benchlist}) and the actual bounds obtained using
our techniques (Table~\ref{fig:comparison}) for the other examples
requiring relaxation. AECS/CD1 delay is not included in
Figure~\ref{fig:graph2}, because although it requires relaxation, the
delay specification (Table~\ref{fig:benchlist}) does not have an upper
bound.

@@

There are two influential factors in determining the relaxation: 1)
execution speed of a target platform and 2) the nature of the program
that runs on this platform. Clock frequency of the processor directly
impacts the value of WCRT. On the other hand, programmer can write a
program such that large data (or control) computations in a single
logical tick increase WCRT of the program.  Assuming we do not change
the architecture of the execution platform (i.e. JOP), the differece
factor can can only be reduced by modifying the benchmark programs such
that the critical paths are broken up using \texttt{pause} constructs,
thereby reducing the WCRT.  Figure~\ref{fig:improv} shows segments of
code from AECS whose WCRT is reduced by such technique.

@@

The result of the upper bound relaxation with reduced WCRT is shown in
Figure~\ref{fig:graph2}.  The upper bounds for delay statements in
AECS/CD2 decreased by 11.81\%, while that for Motor example reduced from
6.42 to 8.05\%. This is expected since the amount of computation
required in control flow is less demanding than data and hence, breaking
the critical paths in Motor did not reduce WCRT as much as AECS/CD2.
Therefore, we can see from the result that the size of the relaxation
depends on how designers write their program as well as type of the
application (i.e. data or control dominated).

>>>>

We have carried out a number of experiments on a set of applications
with real-time constraints. The characteristics of the benchmark set
including required real-time delays and program sizes are shown in
Table~\ref{fig:benchlist}.
{\renewcommand{\arraystretch}{0.9}
%\renewcommand{\texttt}[1] {\begin{lstlisting}[language=Java,morekeywords={delay}] #1 \end{lstlisting}}
\begin{table}[h!]
	\scriptsize
	\centering
	\caption{An overview of the benchmark programs}
	\label{fig:benchlist}
	\scalebox{0.9}{
		\begin{tabular}{llp{0.5cm}p{1cm}}
\toprule
\emph{Benchmark}       & \emph{Delay constructs used in each CD} & \emph{\# of CDs} & 
\emph{Lines of code}        
\\ \midrule
HRTCS                  & \texttt{wait\_inbetween(50.3..200.3 ms)} ($CD1$) & 2                  & 56
\\ \midrule
\multirow{5}{*}{Motor} & \texttt{wait\_exactly(2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
											 &
                                                                                         \texttt{wait\_exactly(1.667
                                                                                           ms)}      &                    &                      \\ 
											 &
                                                                                    \texttt{wait\_exactly(0.05
                                                                                           ms)}       &                    &                      \\ 
											 &
                                                                                    \texttt{wait\_exactly(0.3
                                                                                           ms)}        &                    &                      \\ 
											 &
                                                                                         \texttt{wait\_exactly(0.733
                                                                                           ms)}      &                    &
\\ \midrule
\multirow{3}{*}{Printer}
& \texttt{wait\_inbetween(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
& \texttt{wait\_inbetween(1.2..30.3 ms)} ($CD2$)		  &														&											\\
& \texttt{wait\_inbetween(5.6..100.2 ms)} ($CD2$)     &														&											
\\ \midrule
\multirow{2}{*}{AECS} & \texttt{wait\_atleast(10000 ms)} ($CD1$) 
& \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
& \texttt{wait\_exactly(10000 ms)} ($CD2$) & &
\\ \bottomrule
\end{tabular}
}
\end{table}
}
{\renewcommand{\arraystretch}{0.6}
\begin{table*}[t!]
\centering
	\caption{Actual delays obtained for wait constructs in the
benchmark programs based on their BCRT and WCRT}
\begin{tabular}{ c c l l l l l l }
	\toprule
	\multicolumn{2}{c}{} & \emph{HRTCS} & \emph{Motor} & \emph{Printer/CD1} &
	\emph{Printer/CD2} & \emph{AECS/CD1} &	\emph{AECS/CD2} \\ 
	\midrule
	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0.0396 ms	& 0.0425 ms & 0.0190 ms &
	0.0263 ms & 0.1784 ms & 0.1284 ms\\ 
	\cmidrule(r){1-8}
	\multicolumn{2}{c}{\emph{WCRT}} &0.1152 ms & 0.1096 ms & 0.0647 ms
	& 0.0491 ms & 0.3476 ms & 0.2991 ms\\ 
	\cmidrule(r){1-8}

	\multirow{2}{*}{Delay 1} & M..N & 50.3-200.3 ms	& 2.4-6.2493 ms	  &
	2.7-10.6 ms & 1.2-30.3 ms & 10000-19484.3481 ms& 10000-23278.2752 ms \\
	\cmidrule(r){2-8}
	& d &1506-1788 & 57 & 143-163 & 46-617 & 56062 &77844\\ 
	\cmidrule(r){1-8}
	\multirow{2}{*}{Delay 2} & M..N & N/A & 1.668-4.3855 ms & N/A &
	5.6-100.2 ms & N/A &10000-23278.2752 ms\\ 
	\cmidrule(r){2-8}
	& d & N/A & 40 & N/A & 213-2041 & N/A &53160\\ 
	\cmidrule(r){1-8}
	\multirow{2}{*}{Delay 3} & M..N & N/A & 0.05-0.2193 ms & N/A & N/A & N/A
	&N/A\\ 
	\cmidrule(r){2-8}
	& d & N/A & 2 & N/A & N/A & N/A &N/A\\ 
	\cmidrule(r){1-8}
	\multirow{2}{*}{Delay 4} & M..N & N/A & 0.3-0.8771 ms & N/A & N/A & N/A
	&N/A\\ 
	\cmidrule(r){2-8}
	& d & N/A & 8 & N/A & N/A & N/A &N/A\\ 
	\cmidrule(r){1-8}
	\multirow{2}{*}{Delay 5} & M..N & N/A & 0.733-1.9735 ms & N/A & N/A & N/A
	&N/A\\ 
	\cmidrule(r){2-8}
	& d & N/A & 18 & N/A
	& N/A & N/A &N/A\\ \bottomrule
\end{tabular}
\label{fig:comparison}
\end{table*}
}

HRTCS is the human response system described in
Section~\ref{sec:motivating-example}. Motor is a stepper motor
controller from~\cite{Bourke2009a}, used for producing monochrome images
on paper by correct application of current to a print head (a row of
resistors). In this example, duration of the current applied is
controlled by the feedback logic using real-time waits, which are
expressed using reactive constructs. Printer is an example
from~\cite{Schneider:1999:CRT:555233}, which models a printer and a
spooler in timed CSP. Its operation can be described as follows: as soon
as the spooler receives a print job it passes it to the printer through
a channel where actual printing process take place. This example along
with HRTCS demonstrates the use of non exact wait statements.

AECS (Access and Environmental Control System) is an application that
controls an intelligent room environment described
in~\cite{aecs_ispa}. The system interacts with two external timers
through input/output signals in order to measure how long the entrance
door to the room is being opened to detect intruders.  The program has
been modified such that the wait constructs are used instead of external
timers as shown in Figure~\ref{delaycode}. In the original AECS, the
system needed to trigger external timer by emitting the output signal in
order to wait for a specific amount of time (e.g.
\texttt{TimerTriggered} in Figure~\ref{delaycode:a}). When the timer
expires, it notifies the system through the input signal
\texttt{TimeOut}, which is polled using the \texttt{await}
statement. With the new wait constructs, these timers are not necessary
and waits can be expressed directly in the code as shown in
Figure~\ref{delaycode:b}.

\begin{figure}[h!]
	\vspace{-10pt}
	\centering
  \begin{SubFloat}{\label{delaycode:a}Original program}
      \scriptsize
		\begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,suspend}]
emit TimerTriggered;
trap(T){
  abort(DoorOpened){
    await(TimeOut);
    exit(T); 
  }
}
\end{lstlisting}
  \end{SubFloat}
  \hspace{1cm}%
  \begin{SubFloat}{\label{delaycode:b}Program with wait constructs}
		\begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,wait_atleast,suspend}]
// No emit statement
trap(T){
  abort(DoorOpened){
    wait_atleast (10000 ms);
    exit(T);  
  }
}
\end{lstlisting}
	\end{SubFloat}
\caption{Replacing external timer with wait\_atleast construct}
\label{delaycode}
\end{figure}

@@

HRTCS consists of 2 CDs, but only one with real-time waits. AECS also
has 2 CDs; CD1 consist of a single wait statement, while CD2 has 2 wait
statements. In this case, as explained previously, wait statements
replace the external timers and corresponding interface (I/O)
signals. The system triggers an alarm 10 seconds after detection of an
intruder or generates a beep sound when an entrance door is being opened
for the same amount of time. In the Printer example, spooler and printer
processes are mapped to separate CDs. Spooler CD sends a job to the
Printer CD between 2.7 to 10.6 ms after receiving it from the
environment (e.g. a user). On the other hand, Printer CD notifies the
environment that it starts printing between 1.2 to 30.3 ms once it
receives the job from the Spooler CD. Actual print job is started
between 5.6 to 100.2 ms after the notification. In this example, bounded
wait ensures that both; channel communication and print job should occur
within the specified amount of time. Lastly, Motor is a synchronous
program with 5 exact wait statements that control the amount of current
applied to the stepper motor coils.

@@

Table~\ref{fig:comparison} shows the result of our algorithm applied to
each benchmark program. Note that we allowed the compiler to relax the
upper real-time bounds for \texttt{wait\_exactly} and
\texttt{wait\_inbetween} when it failed to find a feasible solution.
However, a user can prohibit such behavior using compiler switch as
described in Section~\ref{sec:tool-chain-flow}.

As a first step, BCRT and WCRT were calculated for all examples. Given
the real-time \mbox{\texttt{wait\_inbetween (M..N)}} provided by
programmers, required logical delay \textit{d} was determined by
applying Algorithm~\ref{alg:1}. In HRTCS, real-time delay of the range
50.3..200.3 (ms) is only satisfied when \emph{d} is between 1506 and
1788 (as shown previously in Section~\ref{sec:find-logic-delay}). Delay
2.7..10.6 (ms) is satisfied in Printer/CD1 with \emph{d} of 143..163.
Similarly, Printer/CD2 satisfies real-time ranges of 1.2..30.3 (ms) and
5.6..100.2 (ms) when \emph{d} is 46..617 and 213..2041, respectively. It
is then the compiler's job to statically choose one of the values in
this range.

@@

The previous results show that for HRTCS and Printer the compiler was
able to find \textit{d} without relaxing upper bound of wait
statements. Figure~\ref{fig:graph1} shows the accuracy of our results by
comparing differences between programmer specified upper bounds
(Table~\ref{fig:benchlist}) and the actual bounds obtained using our
techniques (Table~\ref{fig:comparison}) for the other examples requiring
relaxation. AECS/CD1 wait specification is not included in
Figure~\ref{fig:graph2}, because although it requires relaxation, the
wait specification (Table~\ref{fig:benchlist}) does not have an upper
bound.

@@

There are two influential factors in determining the relaxation: 1)
execution speed of a target platform and 2) the nature of the program
that runs on this platform. Clock frequency of the processor directly
impacts the value of WCRT. On the other hand, programmer can write a
program such that large data (or control) computations in a single
logical tick increase WCRT of the program.  Assuming we do not change
the architecture of the execution platform (i.e. JOP), the differece
factor can only be reduced by modifying the benchmark programs such that
the critical paths are broken up using \texttt{pause} constructs,
thereby reducing the WCRT.  Figure~\ref{fig:improv} shows segments of
code from AECS whose WCRT is reduced by such technique.

@@

The result of the upper bound relaxation with reduced WCRT is shown in
Figure~\ref{fig:graph2}.  The upper bounds for wait statements in
AECS/CD2 decreased by 11.81\%, while that for Motor example reduced from
6.42 to 8.05\%. This is expected since the amount of computation
required in control flow is less demanding than data and hence, breaking
the critical paths in Motor did not reduce WCRT as much as AECS/CD2.
Therefore, we can see from the result that the size of the relaxation
depends on how designers write their program as well as type of the
application (i.e. data or control dominated).

### 2014-01-27 introduction.tex
Updated the paper

Reactive languages~\cite{gber931,amal10} are a class of programming
languages used for designing and implementing reactive systems, which
continuously respond to input from their environment. These languages
have been successfully used in programming a plethora of systems such as
fly-by-wire in Airbus~\cite{eairbus}, security surveillance
systems~\cite{amal121}, etc. Usually these systems also need to meet
real-time constraints enforced by the environment. Yet, these languages
do not support describing real-time statements as first class language
constructs.  For example, one cannot describe a simple real-time delay
(postponement) of an operation for 0.2 ms. These languages are based on
formal semantics, essential for the formal reasoning about and
verification of correctness of functional properties of the developed
programs, but leave nonfunctional properties such as timing behavior as
an implementation detail~\cite{boldt07}. Arguably, rightly so, because
physical time cannot be incorporated without information about the
underlying execution platform.  But, one can still reason in terms of
logical time. These languages support discrete logical clock rather than
a discrete physical (real-time) clock. The real-time period of the
discrete logical clock (also referred to as a logical tick) is not fixed
and it is determined by the responsiveness of the program to the input
signals. Unlike a discrete physical clock, which has fixed real-time
period, the period of the logical clock is elastic. Although the timing
model with logical clock has worked well for reactive synchronous/GALS
languages in designing discrete systems, there is a need to introduce
real-time since, often, implementation models are in the continuous or
discrete real-time domain~\cite{DBLP:journals/pieee/SifakisTY03}.
Moreover, adding timing capabilities as first class language constructs
create additional opportunity for formally verifying functional and
non-functional requirements before system deployment.

@@

The major motivations for introducing delay mechanism and at the same
time contributions of the paper are: (1) delays allow real-time based
synchronization between concurrent behaviors, (2) delay model relies on
the use of relative instead of absolute real-time, i.e., a delay in
selected time units is counted from the currently executing SystemJ
statement (3) delays allow mixing behaviors with real-time features and
others with only logical time and (4) real-time delay does not affect
the model of logical time in SystemJ program as long as the amount of
delay is within boundaries that can be determined statically by program
analysis. All these as well as semantics of the delay construct are
discussed and illustrated in the remaining part of the paper. Our
contributions can be further refined as follows:
\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
 i.e., non-negative rational numbers.
\item We consider the property of \textit{non
    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
  not always sufficient resources for processes to execute, so
  effectively processes are allocated and scheduled with different
  optimization criteria -- reducing computation time, reducing power
  consumption, etc. This allocation and scheduling is an integral part
  of any language compilation and needs to be considered when
  introducing real-time delays.
% \item Our solution does not change the functional or timing semantics of
%   GALS and its subset synchronous programs, in fact, our solution does
%   not require one to change the mid-end, the back-end or the related
%   optimization phases of the compiler at all.
\item We are the first, to our knowledge, to allow specification of both
  deterministic and non-deterministic real-time in GALS and synchronous
  languages. \textit{Deterministic delay:} is the ability to specify an
  exact \textit{real-time} delay. \textit{Non-deterministic delay:} is
  the ability to specify a range of real-time values for a delay
  statement, e.g., delay statement in Figure~\ref{fig:1}.
\end{enumerate*}

The rest of the paper is arranged as follows:
Section~\ref{sec:background} gives the background on SystemJ and related
techniques required for the rest of the
paper. Section~\ref{sec:intr-real-time} introduces real-time delays in
the GALS paradigm and describes their compilation into logical
time. Section~\ref{sec:experimental-results} gives the experimental
results for a set of real-time
applications. Section~\ref{sec:related-work} positions this work in
relation to other approaches to introducing real-time. Finally, we
conclude in Section~\ref{sec:concl-future-work}.

>>>>

Reactive languages~\cite{gber931,amal10} are a class of programming
languages used for designing and implementing reactive systems, which
continuously respond to input from their environment. These languages
have been successfully used in programming a plethora of systems such as
fly-by-wire in Airbus~\cite{eairbus}, security surveillance
systems~\cite{amal121}, etc. Usually these systems also need to meet
real-time constraints imposed by the environment. Yet, these languages
do not support describing real-time statements as first class language
constructs.  For example, one cannot describe a simple real-time delay
(postponement) of an operation for 0.2 ms. These languages are based on
formal semantics, essential for the formal reasoning about and
verification of correctness of functional properties of the developed
programs, but leave nonfunctional properties such as timing behavior as
an implementation detail~\cite{boldt07}. Arguably, rightly so, because
physical time cannot be incorporated without information about the
underlying execution platform.  But, one can still reason in terms of
logical time. These languages support discrete logical clock rather than
a discrete physical (real-time) clock. The real-time period of the
discrete logical clock (also referred to as a logical tick) is not fixed
and it is determined by the responsiveness of the program to the input
signals. Unlike a discrete physical clock, which has fixed real-time
period, the period of the logical clock is elastic. Although the timing
model with logical clock has worked well for reactive synchronous/GALS
languages in designing discrete systems, there is a need to introduce
real-time since, often, implementation models are in the continuous or
discrete real-time domain~\cite{DBLP:journals/pieee/SifakisTY03}.
Moreover, adding timing capabilities as first class language constructs
create additional opportunity for formally verifying functional and
non-functional requirements before system deployment.

@@

The major motivations for introducing these real-time \textit{wait}
mechanisms and at the same time \textbf{contributions} of the paper are:
(1) these waiting mechanisms allow real-time based synchronization
between concurrent behaviors, (2) the waiting model relies on the use of
relative instead of absolute real-time, i.e., a wait in selected time
units is counted from the currently executing SystemJ statement, (3)
these waiting mechanisms allow mixing behaviors with real-time features
and others with only logical time and (4) real-time wait does not affect
the model of logical time in SystemJ program as long as the amount of
delay is within boundaries that can be determined statically by program
analysis. All these as well as semantics of the delay construct are
discussed and illustrated in the rest of the paper. Our contributions
can be further refined as follows:
\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
 i.e., non-negative rational numbers.
\item We consider the property of \textit{non
    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
  not always sufficient resources for processes to execute, so
  effectively processes are allocated and scheduled with different
  optimization criteria -- reducing computation time, reducing power
  consumption, etc. This allocation and scheduling is an integral part
  of any language compilation and needs to be considered when
  introducing real-time delays.
\item We are the first, to our knowledge, to allow specification of
  both; exact and non-exact real-time in GALS and synchronous programs.
\item We elucidate upon a key insight relating real-time and
  logical-time -- \textit{the resolution of any real-time statement is
    dependent on the worst case logical tick time} -- an aspect that
  seems to have escaped in the current literature of real-time in
  synchronous/GALS languages.
\end{enumerate*}

The rest of the paper is arranged as follows:
Section~\ref{sec:background} gives the background on SystemJ and related
techniques required for the rest of the
paper. Section~\ref{sec:intr-real-time} introduces real-time mechanisms
in the GALS paradigm and describes their compilation into logical
time. Section~\ref{sec:experimental-results} gives the experimental
results for a set of real-time
applications. Section~\ref{sec:related-work} positions this work in
relation to other approaches to introducing real-time. Finally, we
conclude in Section~\ref{sec:concl-future-work}.
